SPARQL on the Open, Decentralised Web

### 2016-07-30 dissertation/introduction.tex
Add label
Fix mssing
software contributions

\subsection{Interacting With Data}

@@

\begin{example}{A triple pattern}
\begin{verbatim}
?subject a ?class .
\end{verbatim}
\end{example}
This triple pattern will match the two triples 
\begin{example}{Triples matched by triple pattern}
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
from the above example (prefixes omitted for brevity). Frameworks in
popular programming languages commonly have a method that implements
this.

@@

However, even though the result does not depend on the order of
execution of a BGP, the time it takes to find the result often does. For
example, consider the following BGP:
\begin{example}{A Basic Graph Pattern}
\begin{verbatim}
?subject foaf:name ?name ;
         foaf:workplaceHomepage <http://example.com/> ;
         foaf:made ?contribution .
\end{verbatim}
\end{example}
In most cases, if the triple pattern with the \rdfterm{foaf:name}
predicate is evaluated first, it is likely to match many resources,
since nearly everything may have a name, and therefore, it does not
reduce the number of statements that needs to be searched when
matching the next triple pattern by much. Instead, the
\rdfterm{foaf:workplaceHomepage} may be matched first, as we may guess
the fact that the object is given will be more restrictive, but that
too fails if everyone in the database works for the same organisation
and has used this predicate. It is the query planner's task to choose
between equivalent plans to ensure that the most efficient evaluation
is chosen. As is clear from this example, it is often insufficient to
rely on such heuristics, and therefore, it is common to rely on a
statistical digest in the query planner. Still, the task of query
planning remains a very complex one, and often, insufficient
statistics is available, so that planning is done using heuristics or
based on assumptions that do not hold. These are only some of the
concerns that affect performance of the overall query engine.

@@

>>>>

\subsection{Interacting With Data}\label{sec:introinteract}

@@

\begin{example}{A triple pattern}
\begin{verbatim}
?subject a ?class .
\end{verbatim}
\end{example}
This triple pattern will match the two triples 
\begin{example}{Triples matched by triple pattern}
\begin{verbatim}
<http://www.kjetil.kjernsmo.net/foaf#me> a foaf:Person .
<http://folk.uio.no/kjekje/2013/iswc.pdf> a foaf:Document .
\end{verbatim}
\end{example}
from the above example (prefixes omitted for brevity). Frameworks in
popular programming languages commonly have a method that implements
this.

@@

However, even though the result does not depend on the order of
execution of a BGP, the time it takes to find the result often does. For
example, consider the following BGP:
\begin{example}{A Basic Graph Pattern}\label{ex:bgp}
\begin{verbatim}
?subject foaf:name ?name ;
         foaf:workplaceHomepage <http://example.com/> ;
         foaf:made ?contribution .
\end{verbatim}
\end{example}
In most cases, if the triple pattern with the \rdfterm{foaf:name}
predicate is evaluated first, it is likely to match many resources,
since nearly everything may have a name, and therefore, it does not
reduce the number of statements that needs to be searched when
matching the next triple pattern by much. Instead, the
\rdfterm{foaf:workplaceHomepage} may be matched first, as we may guess
the fact that the object is given will be more restrictive, but that
too fails if everyone in the database works for the same organisation
and has used this predicate. It is the query planner's task to choose
between equivalent plans to ensure that the most efficient evaluation
is chosen. As is clear from this example, it is often insufficient to
rely on such heuristics, and therefore, it is common to rely on a
statistical digest in the query planner. Still, the task of query
planning remains a very complex one, and often, insufficient
statistics is available, so that planning is done using heuristics or
based on assumptions that do not hold. These are only some of the
concerns that affect performance of the overall query engine.

@@

\subsection{Software}

\begin{enumerate}
\item 5 Free Software packages that have been accepted into the Debian
  GNU/Linux operating system.
\item Further 5 packages that already were accepted in Debian have been
  enhanced in the course of this work.
\item Accepted contributions to the code base of another 5 external projects.
\end{enumerate}
